Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-24 main_new.tex
updates

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a,b}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{c}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{d}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}

@@

\address{Department of Engineering Cybernetics,\printead{e4}}
\affiliation[e]{Department of Engineering Cybernetics, NTNU}
\affiliation[f]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

@@

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task for
  modern marine management and science. Faced with limited resources to
  understand highly dynamic processes in the water-column, the
  combination of statistics and autonomous
  robotics provides opportunities for studying aspects related to
  experimental design for effective characterization of oceanographic
  phenomena. In this work we develop methods for efficient spatial
  sampling applied to the mapping of coastal processes by
  providing informative descriptions of spatial characteristics of ocean
  phenomena. Specifically, we define a design criterion based on
  reducing uncertainty in the joint excursions of random
  fields, and derive analytical expressions for the expected Bernoulli
  variance reduction in the excursion set of spatial temperature and
  salinity variables. We show how these can be applicable in static
  designs and sequential designs, building on myopic and look-ahead
  sampling strategies. We use simulations to study the properties and
  differences between state-of-the-art approaches. Results from field
  deployments with an autonomous underwater vehicle are also presented,
  for a case study mapping the boundary of a river plume. The results
  demonstrate the ability of the method to effectively inform and
  execute data-driven environmental sampling.

@@

An important synergy underlining this work is the potential of robotic sampling to adapt survey plans based on data and statistical inference,
in accordance with the sense-plan-act loop
(Fig. \ref{fig:sense-plan-act}). We will therefore present a few adaptive strategies
for robotic sampling that builds on the notion of multivariate ES. Simulations, analysis, and  results from full-scale trials are presented.

@@

The optimal sequential design not only considers the best current AUV
grid node, but also what data gathered at this node would lead to for
future sampling at successive stages.  Let $d^{j,s}$ denote design
number $j$ at stage $s$ of a sequential survey. If this design is
selected, data $\by^{j,s}$ will be gathered. The optimal path
selection situation is then depicted in Figure \ref{fig:PathSelOpt},
\begin{figure}[h!]
\centering
\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
\end{figure}
where design choices are indicated by squares while data realizations are indicated by circles. 
%terms this optimal design is defined as
%\begin{equation}\label{opt_crit}
%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
%\end{equation}
%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
The mathematical expression for the optimal design then involves a
series of intermixed maximizations over designs and integrals over
data.  In practice the optimal solution is not tractable because of
the enormous growth over stages, see
e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
Instead, we outline heuristic strategies. In terms of notation, let
$\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote
the data gathered until stage $s-1$, using a selected design
$d^{j,r}$, $r=1,\ldots,s-1$.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
\caption{The equilateral waypoint graph used to discretize the
  trajectory choices over the $31\times31$ grid used to discretize the GP.}
\label{fig:wp_graph}
\end{figure}

@@

\begin{figure}[!h] 
 \centering 
\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
\caption{Data analysis from a preliminary trial experiment using the
  AUV. Left: Residual plot of temperature and salinity along with
  Gaussian contours. Middle: Empirical CDF of the quadratic form of
  the residuals along with the theoretical CDF of the $\chi^2$
  distribution with two degrees of freedom. Right: Empirical variogram
  of the salinity and temperature data.} \label{fig:parest}
\end{figure}

@@

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater Vehicle (LAUV) platform (Harald) for upper water-
column exploration used in the experiment.}
\label{fig:lauv}
\end{figure}

>>>>

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a,b}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{c}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{d}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{b,e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}

@@

\address{Department of Engineering Cybernetics,\printead{e4}}
\affiliation[e]{Department of Engineering Cybernetics, NTNU}
\affiliation[f]{Underwater Systems and Technology Laboratory (LSTS),
  Faculty of Engineering, University of Porto, Portugal.}

@@

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for modern marine management and science. Faced with limited
  resources to understand highly dynamic processes in the
  water-column, the combination of statistics and autonomous robotics
  provides opportunities for studying aspects related to experimental
  design for effective characterization of oceanographic phenomena. In
  this work, we develop methods for efficient spatial sampling applied
  to the mapping of coastal processes by providing informative
  descriptions of spatial characteristics of ocean
  phenomena. Specifically, we define a design criterion based on
  reducing uncertainty in the joint excursions of random fields, and
  derive analytical expressions for the expected Bernoulli variance
  reduction in the excursion set of spatial temperature and salinity
  variables. We show how these can be applicable in static designs and
  sequential designs, building on myopic and look-ahead sampling
  strategies. We use simulations to study the properties and
  differences between state-of-the-art approaches. Results from field
  deployments with an autonomous underwater vehicle are also
  presented, for a case study mapping the boundary of a river
  plume. The results demonstrate the ability of the method to
  effectively inform and execute data-driven environmental sampling.

@@

An important synergy underlining this work is the connection to
robotic sampling to adapt survey plans based on data and statistical
inference, in accordance with the sense-plan-act loop
(Fig. \ref{fig:sense-plan-act}). To do so we build on the notion of
multivariate ES and simulations, analysis, and results from full-scale
robotic trials are presented.

@@

The optimal sequential design not only considers the best current AUV
grid node, but also what data gathered at this node would lead to for
future sampling at successive stages.  Let $d^{j,s}$ denote design
number $j$ at stage $s$ of a sequential survey. If this design is
selected, data $\by^{j,s}$ will be gathered. The optimal path
selection situation is then depicted in Figure \ref{fig:PathSelOpt},
\begin{figure}[!b]
\centering
\includegraphics[width=0.85\textwidth]{Figures/sequent_select.pdf}
\caption{Optimal sequential path selection.}\label{fig:PathSelOpt}
\end{figure}
where design choices are indicated by squares while data realizations are indicated by circles. 
%terms this optimal design is defined as
%\begin{equation}\label{opt_crit}
%    \bd^* = \mbox{argmin}_{d^{j,1}} \left\{ \int_{\by^{j,1}} \mbox{argmin}_{d^{j,2}} \left\{ \int_{\by^{j,2}} \ldots \pi(\by^{j,2}|\by^{j,1}) d\by^{j,2} \right\} \pi(\by^{j,1}) d \by^{j,1} \right\},
%\end{equation}
%where $\ldots$ represents the expected variance reduction in the ES under the continued optimal design, which will depend on the data at earlier stages. 
The mathematical expression for the optimal design then involves a
series of intermixed maximizations over designs and integrals over
data.  In practice the optimal solution is not tractable because of
the enormous growth over stages, see
e.g. \cite{sucar2015probabilistic} and \cite{powell2016perspectives}.
Instead, we outline heuristic strategies. In terms of notation, let
$\mathcal{Y}^{s-1} =\{d^{j,r}, \by^{j,r} ; r=1,\ldots,s-1 \}$ denote
the data gathered until stage $s-1$, using a selected design
$d^{j,r}$, $r=1,\ldots,s-1$.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.50\textwidth]{Figures/sim/wp_graph_paper.pdf}
\caption{The equilateral waypoint graph used to discretize the
  trajectory choices over the $31\times31$ grid used to discretize the
  Gaussian Process.}
\label{fig:wp_graph}
\end{figure}

@@

\begin{figure}[!b] 
 \centering 
\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
\caption{Data analysis from a preliminary trial experiment using the
  AUV. Left: Residual plot of temperature and salinity along with
  Gaussian contours. Middle: Empirical CDF of the quadratic form of
  the residuals along with the theoretical CDF of the $\chi^2$
  distribution with two degrees of freedom. Right: Empirical variogram
  of the salinity and temperature data.} \label{fig:parest}
\end{figure}

@@

\begin{figure}[!h] 
\centering 
\includegraphics[width=0.98\textwidth]{Figures/harald.jpg}
\caption{The commercially available Light Autonomous Underwater
  Vehicle (LAUV) platform for upper water- column exploration used in
  the experiment.}
\label{fig:lauv}
\end{figure}
